---
title: "NYPD_Shooting_Incident_Data_historic"
author: "C. Chen"
date: "1/26/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description:
2 visualizations are done with the NYPD Shooting Incident Data. The variables used are: the BORO, OCCUR_DATE and MURDER_FLAG. The data was transformed to categorize by borough and 2 graphs are created. One showing the Shootings per borough and the other one showing the murders per borough. 

Later another data subset is created with the data from the Staten Island borough and a linear prediction is made comparing shootings and murders. As a result we see that the prediction is a linear trend which proves that there's a relationship between the shootings and murders but there's definitely other factors that need to included in order to have a better prediction due to the outliers.

Some Bias is that we are not taking into consideration the population proportions to compare the boroughs. Another bias can be racial because you can get into a conclusion but I would like to know more about the social status and environments related to each incident. I noticed that there's a lot of missing data for perpetrator. I wanted to do an analysis comparing the victims and the perpetrator but noticed there was a lot of missing data.  
```{r get_jhu_data}
library(tidyverse)
url_in<-"https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
shooting_incident_data<-read_csv(url_in)
shooting_incident_data<-subset (shooting_incident_data, select = -c(LOCATION_DESC,INCIDENT_KEY,OCCUR_TIME,PRECINCT,PERP_AGE_GROUP,PERP_SEX, PERP_RACE,X_COORD_CD, Y_COORD_CD, Latitude, Longitude, Lon_Lat,JURISDICTION_CODE))
colnames(shooting_incident_data)[colnames(shooting_incident_data) == "BORO"] <- "BOROUGH"
colnames(shooting_incident_data)[colnames(shooting_incident_data) == "VIC_AGE_GROUP"] <- "VICTIM_AGE"
colnames(shooting_incident_data)[colnames(shooting_incident_data) == "VIC_SEX"] <- "VICTIM_SEX"
colnames(shooting_incident_data)[colnames(shooting_incident_data) == "VIC_RACE"] <- "VICTIM_RACE"
library(lubridate)
shooting_incident_data <- shooting_incident_data %>%
    mutate(DATE=mdy(OCCUR_DATE),MURDER=as.integer(shooting_incident_data$STATISTICAL_MURDER_FLAG),countOccurence=1)
summary(shooting_incident_data)
shootings_by_boro <-shooting_incident_data%>%
  group_by(BOROUGH)%>%
  summarize(MURDERS = sum(MURDER),SHOOTINGS = sum(countOccurence)) %>%
  select(BOROUGH,MURDERS,SHOOTINGS)%>%
  ungroup()
ggplot(shootings_by_boro, aes(x = BOROUGH, y = SHOOTINGS, fill = SHOOTINGS)) + 
     geom_bar(stat = "identity") +
     geom_text(aes(label = SHOOTINGS), vjust = 0) +ggtitle("Shootings per Borough")+
     theme(plot.title=element_text(hjust=0.5))
ggplot(shootings_by_boro, aes(x = BOROUGH, y = MURDERS, fill = MURDERS)) + 
     geom_bar(stat = "identity") +
     geom_text(aes(label = MURDERS), vjust = 0) +ggtitle("Murders per Borough")+
     theme(plot.title=element_text(hjust=0.5))
shootings_by_dates <-shooting_incident_data%>%
  group_by(DATE,BOROUGH)%>%
  summarize(MURDERS = sum(MURDER),SHOOTINGS = sum(countOccurence)) %>%
  select(BOROUGH,MURDERS,SHOOTINGS)%>%
  ungroup()
boro <-"STATEN ISLAND" 
shootings_by_dates %>%
  filter(BOROUGH == boro) %>%
  ggplot(aes(x = DATE, y = SHOOTINGS))+
  geom_line(aes(color = "SHOOTINGS")) +
  geom_point(aes(color = "SHOOTINGS"))+
  geom_line(aes(y= MURDERS, color = "MURDERS")) +
  geom_point(aes(y=MURDERS, color = "MURDERS")) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle =90,hjust=1)) +
  labs(title = str_c("Shootings vs Murders in ", boro), y=NULL)
shootings_by_Staten_Island<-shootings_by_dates %>% 
     filter(BOROUGH == "STATEN ISLAND") %>%
     select(BOROUGH,MURDERS,SHOOTINGS,DATE)%>%
     ungroup()
mod<-lm( MURDERS ~ SHOOTINGS,data=shootings_by_Staten_Island)
summary(mod)
shootings_by_Staten_Island_with_pred<-shootings_by_Staten_Island%>%
  mutate(pred=predict(mod))
shootings_by_Staten_Island_with_pred %>% ggplot() +
 geom_point(aes(x = SHOOTINGS, y = MURDERS ), color ="blue")+
 geom_point(aes(x = SHOOTINGS, y = pred), color = "red")
```